Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2018-02-08 paper-ieee-trans.tex
remove footnotes

\author{E. Ronchieri,
        P. Orviz Fernandez,
        M. David,
	D. C. Duma,
        J. Gomes,
        D. Salomoni
\thanks{D. C. Duma, E. Ronchieri, and D. Salomoni are with INFN - CNAF, Bologna, Italy.}
\thanks{P. Orviz Fernandez, is with CSIC, Santander, Spain, (e-mail: orviz@ifca.unican.es)}
\thanks{M. David, and J. Gomes are with LIP, Lisbon, Portugal.}%
}

@@

\begin{IEEEkeywords}
Software Reliability, Quality Assurance, Software Metrics, Software Testing
Techniques
\end{IEEEkeywords}

@@

\IEEEPARstart{D}uring the last 15 years the provision of an European e-Infrastructure,
which supports an unified access to large-scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan-European research collaborations. One of the well-known examples is the Worldwide 
Large Hadron Collider (LHC) Computing Grid infrastructure, powered 
by the particle physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large 
Hadron Collider particle accelerator. A significant number of innovation projects and
technology transfer activities -- embraced by the 5th, 6th and 7th Framework Programmes,
and ultimately by the Horizon 2020 programme\footnote{HORIZON 2020 - The EU Framework
Programme for Research and Innovation https://ec.europa.eu/programmes/horizon2020/} -- are being held to coordinate and 
develop distributed solutions that provide a seamless and transparent access to those
computing and data resources, available for the researchers.

@@

\item \textit{Automated Deployment}.
In order to lower the barriers of software adoption, the SQA criteria requires the automated 
deployment of the products delivered as part of the catalogue. Automation in this context is 
tackled using a configuration management tool. These tools allow to have an 
\textit{infrastructure as code} (IaC), managing the component's deployment through declarative
definitions. The definition files appear as additional source of documentation, 
self-documenting code, as they sequentially guide the component's deployment process in multiple
platforms. In this regard, {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such 
as Ansible \cite{indigo-ansible} and Puppet \cite{indigo-puppet}. A representative example of 
such contribution are the \textit{50 roles} being developed from scratch, currently hosted in
the Ansible Galaxy portal \footnote{https://galaxy.ansible.com/indigo-dc/}. Figure 
\ref{fig:fig_confman} shows the number of products that offer an automated means for deployment.
It shows an increase after the {\sl INDIGO--1} release especially in the previous weeks of each
of the two major releases.
\end{itemize}

@@

@@

>>>>

\author{E. Ronchieri,
        P. Orviz Fernandez,
        M. David,
	D. C. Duma,
        J. Gomes,
        D. Salomoni
\thanks{D. C. Duma, E. Ronchieri, and D. Salomoni are with INFN - CNAF, Bologna, Italy.}
\thanks{P. Orviz Fernandez, is with CSIC, Santander, Spain.}
\thanks{M. David, and J. Gomes are with LIP, Lisbon, Portugal.}%
}

@@

\begin{IEEEkeywords}
Software Reliability, Quality Assurance, Software Metrics, Software Testing
Techniques, DevOps
\end{IEEEkeywords}

@@

\IEEEPARstart{D}uring the last 15 years the provision of an European e-Infrastructure,
which supports an unified access to large-scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan-European research collaborations. One of the well-known examples is the Worldwide 
Large Hadron Collider (LHC) Computing Grid infrastructure, powered 
by the particle physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large 
Hadron Collider particle accelerator. A significant number of innovation projects and
technology transfer activities -- embraced by the 5th, 6th and 7th Framework Programmes,
and ultimately by the Horizon 2020 programme \cite{h2020} -- are being held to coordinate and 
develop distributed solutions that provide a seamless and transparent access to those
computing and data resources, available for the researchers.

@@

\item \textit{Automated Deployment}.
In order to lower the barriers of software adoption, the SQA criteria requires the automated 
deployment of the products delivered as part of the catalogue. Automation in this context is 
tackled using a configuration management tool. These tools allow to have an 
\textit{infrastructure as code} (IaC), managing the component's deployment through declarative
definitions. The definition files appear as additional source of documentation, 
self-documenting code, as they sequentially guide the component's deployment process in multiple
platforms. In this regard, {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such 
as Ansible \cite{indigo-ansible} and Puppet \cite{indigo-puppet}. A representative example of 
such contribution are the \textit{50 roles} being developed from scratch, currently hosted in
the Ansible Galaxy portal \cite{indigo-galaxy}. Figure 
\ref{fig:fig_confman} shows the number of products that offer an automated means for deployment.
It shows an increase after the {\sl INDIGO--1} release especially in the previous weeks of each
of the two major releases.
\end{itemize}

@@

\bibitem{h2020}
\emph{HORIZON 2020}, The EU Framework Programme for Research and Innovation [Online]. Available: https://ec.europa.eu/programmes/horizon2020/

@@

\bibitem{indigo-galaxy}
\emph{INDIGO--DataCloud Ansible Galaxy repository},
https://galaxy.ansible.com/indigo-dc/
